---
title: "Prediction Model For Movement Execution"
author: "mml"
date: "Sunday, January 25, 2015"
output: html_document
---
##The dumbell lift, are you doing it right?
###Introduction
In this project, your goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The five different ways are : 

-- exactly according to the specification (Class A)  
-- throwing the elbows to the front (Class B)  
-- lifting the dumbbell only halfway (Class C)  
-- lowering the dumbbell only halfway (Class D)  
-- throwing the hips to the front (Class E)  

The goal of the project is to predict the manner in which they did the exercise according to their respective letter *A* *B* *C* *D* or *E*. This is the "classe" variable in the training set. We can use any of the other variables to predict with. We should create a report describing how we built our model, how we used cross validation, what we think the expected out of sample error is, and why we made the choices we did. We will therefore use our prediction model to predict 20 different test cases. 

training data set for this project is ``https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv``  
test data set for this project is ``https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv``  
more info is avalaible at ``http://groupware.les.inf.puc-rio.br/har``  

###Pre-predictive work
####Cross-validation:
1. Use the training set
2. Split it into training/test sets
3. Build a model on the training set
4. Evaluate on the test set
5. Repeat and average estimated errors

subtraining data: 75% of the training data set  
subtest data: 25% of the training data  
The models will fit on the subtraining data and tested on subset.    
We will test on the real test set once the most/best predictive model is found.  
  
The expected out of sample error will be 1-accuracy found from the cross-validation data  
set because ``classe`` is unordered (although we could have put an order to it) and is a  
factor variable.

###Code and Predictive model
#####Libraries, seed and data
```{r, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(e1071)
set.seed(1210)
train_set <-read.csv("malearntrain.csv", na.strings=c("NA","#DIV/0!",
                                                                                      ""))
test_set <- read.csv("machlearntest.csv", na.strings=c("NA","#DIV/0!", "
"))
```
#####Cleaning data and exploration
```{r, echo=TRUE, results="hide", warning=FALSE,message=FALSE}
dim(train_set)
dim(test_set)
head(train_set)
head(test_set)
##delete col with missing values
train_set<-train_set[,colSums(is.na(train_set)) == 0]
test_set <-test_set[,colSums(is.na(test_set)) == 0]
train_set <-train_set[,-c(1:7)]
test_set <-test_set[,-c(1:7)]
##check the new data set
dim(train_set)
dim(test_set)
head(train_set)
head(test_set)
```

#####Partitioning
```{r, echo=TRUE}
subsamples <- createDataPartition(y=train_set$classe, p=0.75, list=FALSE)
subTraining <- train_set[subsamples, ]
subTesting <- train_set[-subsamples, ]
dim(subTraining)
dim(subTesting)
head(subTraining)
head(subTesting)
#overview
plot(subTraining$classe, col="red", main="Bar Plot of classe levels", xlab="classe levels", ylab="Frequency")
```

#####Modeling

######First Model
```{r, echo=TRUE}
mod1 <- rpart(classe ~ ., data=subTraining, method="class")
# Predicting:
pred1 <- predict(mod1, subTesting, type = "class")
# Plot of the Decision Tree
rpart.plot(mod1, main="classe Tree", extra=102, under=TRUE, faclen=0)
##test results on subTesting data
confusionMatrix(pred1, subTesting$classe)
```

######Second Model
```{r, echo=TRUE}
mod2 <- randomForest(classe ~. , data=subTraining, method="class")
# Predicting:
pred2 <- predict(mod2, subTesting, type = "class")
# Test results on subTesting data set:
confusionMatrix(pred2, subTesting$classe)
```

#####Conclusion:
*Random Forest is performing better than model 1.*

####Submission
```{r, echo=TRUE}
pred_subm <- predict(mod2, test_set, type="class")
pred_subm

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(pred_subm)
```
###References
-- 1. http://groupware.les.inf.puc-rio.br/har  
-- 2.Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.    
-- 3. Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.